Comparing human and model-based forecasts of COVID-19 in Germany and Poland

### 2022-01-11 manuscript/manuscript.tex
update manuscript formatting, add author summary

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Comparing human and model-based forecasts of COVID-19 in Germany and Poland},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin=1in]{geometry}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\usepackage{float}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi

@@

\emph{Nikos I. Bosse, Sam Abbott, Johannes Bracher, Habakuk Hain, Billy J. Quilty, Mark Jit, Centre for the Mathematical Modelling of Infectious Diseases COVID-19 Working Group, Edwin van Leeuwen, Anne Cori, Sebastian Funk}

@@

@@

Infectious disease modelling has a long tradition and has helped inform public health decisions both through scenario modelling, as well as actual forecasts of (among others) influenza (Biggerstaff et al., 2016; e.g. McGowan et al., 2019; Reich, Brooks, et al., 2019; Shaman \& Karspeck, 2012), dengue fever (Colón-González et al., 2021; e.g. Johansson et al., 2019; Yamana et al., 2016), ebola (Funk et al., 2019; e.g. Viboud et al., 2018), chikungunya (e.g. Del Valle et al., 2018; Farrow et al., 2017) and now COVID-19 (Bracher, Wolffram, Deuschel, Görgen, Ketterer, Ullrich, Abbott, Barbarossa, Bertsimas, Bhatia, Bodych, Bosse, Burgard, Castro, et al., 2021; Bracher, Wolffram, Deuschel, Görgen, Ketterer, Ullrich, Abbott, Barbarossa, Bertsimas, Bhatia, Bodych, Bosse, Burgard, Fiedler, et al., 2021; Cramer et al., 2021; Cramer et al., 2020; European Covid-19 Forecast Hub, 2021; e.g. Funk et al., 2020). Applications of epidemiological models differ in the way they make statements about the future. Forecasts aim to predict the future as it will occur, while scenario modelling and projections aim to represent what the future could look like under certain scenario assumptions or if conditions stayed the same as they were in the past. Forecasts can be judged by comparing them against observed data. Since it is much harder to fairly assess the accuracy and usefulness of projections and scenario modelling in the same way, this work focuses on forecasts, which represent only a subset of all epidemiological modelling.

Since March 2020, forecasts of COVID-19 from multiple teams have been collected, aggregated and compared by Forecast Hubs such as the US Forecast Hub (Cramer et al., 2021; Cramer et al., 2020), the German and Polish Forecast Hub (Bracher, Wolffram, Deuschel, Görgen, Ketterer, Ullrich, Abbott, Barbarossa, Bertsimas, Bhatia, Bodych, Bosse, Burgard, Castro, et al., 2021; Bracher, Wolffram, Deuschel, Görgen, Ketterer, Ullrich, Abbott, Barbarossa, Bertsimas, Bhatia, Bodych, Bosse, Burgard, Fiedler, et al., 2021) and the European Forecast Hub (European Covid-19 Forecast Hub, 2021). Often, different individual forecasts are combined into a single forecast, e.g.~by taking the mean or median of all forecasts. These ensemble forecasts usually tend to perform better and more consistently than individual forecasts (see e.g. Yamana et al. (2016); Reich, McGowan, et al. (2019)).

@@

Previous work has examined such direct human forecasts in various contexts, such as geopolitics (Atanasov et al., 2016; Tetlock et al., 2014), meta-science (Hoogeveen et al., 2020; ReplicationMarkets, 2020), sports (Servan-Schreiber et al., 2004) and epidemiology (Farrow et al., 2017; McAndrew \& Reich, 2020; Recchia et al., 2021). Several prediction platforms (CSET Foretell, 2021; Hypermind, 2021; Metaculus, 2020) and prediction markets (PredictIt, 2021) have been created to collate expert and non-expert predictions. However, with the notable exception of Farrow et al. (2017), these forecasts were not designed to be evaluated alongside model-based forecasts and usually follow their own (often binary) prediction formats. Direct human forecasts may be able to take into account insights and relationships between variables which are hard to specify using epidemiological models. However, it is not entirely clear in which situations human forecasts perform well or badly. For example, Farrow et al. (2017) found that humans could outperform computer models at predicting the 2014/15 and 2015/16 flu season in the US, a setting where the disease was well known and information about previous seasons was available. However, humans tended to do slightly worse at predicting the 2014/15 outbreak of chikungunya in the Americas, a disease previously largely unobserved and unknown in these regions at the time.

In this study, we analyse the performance of direct human forecasts relative to model-based forecasts and discuss the added benefit of epidemiological modelling over human judgement alone. As a case study, we use different forecasts, involving varying degrees of human intervention, which we submitted in real time to the German and Polish Forecast Hub. In contrast to Farrow et al. (2017) we elicited not only point predictions, but full predictive distributions (``probabilistic forecasts'', see e.g. Held et al. (2017)) from participants. This allows us to compare not only predictive accuracy, but also how well human forecasters and model-based forecasts were able to quantify forecast uncertainty.

@@

We created and submitted the following forecasts to the German and Polish Forecast Hub: 1) a direct human forecast (henceforth called ``crowd forecast''), elicited from participants through a web application (Bosse, Abbott, EpiForecasts, et al., 2020b) and 2) two semi-mechanistic model-based forecasts (``renewal model'' and ``convolution model'') informed by basic assumptions about COVID-19 epidemiology. While the two semi-mechanistic forecasts were necessarily shaped by our implicit assumptions and decisions, they were designed such as to minimise the amount of human intervention involved. For example, we refrained from adjusting model outputs or refining the models based on past performance. Forecasts were created in real time over a period of 21 weeks from October 12th 2020 until March 1st 2021 and submitted to the German and Polish Forecast hub (Bracher, Wolffram, Deuschel, Görgen, Ketterer, Ullrich, Abbott, Barbarossa, Bertsimas, Bhatia, Bodych, Bosse, Burgard, Castro, et al., 2021; Bracher, Wolffram, Deuschel, Görgen, Ketterer, Ullrich, Abbott, Barbarossa, Bertsimas, Bhatia, Bodych, Bosse, Burgard, Fiedler, et al., 2021). All code and tools necessary to generate the forecasts and make a forecast submission are available in the \texttt{covid.german.forecasts} R package (Bosse, Abbott, EpiForecasts, et al., 2020a). This repository also contains a record of all forecasts submitted to the German and Polish Forecast Hub. Forecasts were evaluated using a variety of scoring metrics and compared among each other and against an ensemble of all other models submitted to the German and Polish Forecast Hub.

@@

The German and Polish Forecast Hub (now mostly merged into the European Covid-19 Forecast Hub (2021)) elicits predictions for various COVID-19 related forecast targets from different research groups every week. Forecasts had to be made every Monday (with submissions allowed until Tuesday 3pm) and were permitted to use any data that was available by Monday 11.59pm. We submitted forecasts for incident and cumulative weekly reported numbers of cases of and deaths from COVID-19 on a national level in Germany and Poland over a one to four week forecast horizon. Forecasts were submitted on Mondays, but weeks were defined as ending on a Saturday (and starting on Sunday), meaning that forecast horizons were in fact 5, 12, 19 and 26 days. Submissions were required in a quantile-based format with 23 quantiles of each output measure at levels 0.01, 0.025, 0.05, 0.10, 0.15, \ldots, 0.95, 0.975, 0.99. Forecasts submitted to the Forecast Hub were combined into different ensembles every week, with the median ensemble (i.e., the \(\alpha\)-quantile of the ensemble is given by the median of all submitted \(\alpha\)-quantiles) being the default ensemble shown on all official Forecast hub visualisations (\url{https://kitmetricslab.github.io/forecasthub/forecast}).

Data on daily reported test positive cases and deaths linked to COVID-19 were provided by the organisers of the German and Polish Forecast hub. Until December 14th, 2020, these data were sourced from the European Centre for Disease Control (ECDC, 2020). After ECDC stopped publishing daily data, observations were sourced from the Robert Koch Institute (RKI) and the Polish Ministry of Health for the remainder of the submission period (RKI, 2021). These data are subject to reporting artefacts, (such as for example delayed case reporting in Poland on the 24th November, Forsal.pl (2020)), changes in reporting over time, and variation in testing regimes (for example in Germany from the 11th of November on, Ärzteblatt (2020)).

@@

Our crowd forecasts were created as an ensemble of forecasts made by individual participants every week through a web application (\url{https://cmmid-lshtm.shinyapps.io/crowd-forecast/}). Weekly forecasts had to be submitted before Tuesday 12pm every week, but participants were asked to only use any information or data that was already available by Monday night. The application was built using the \texttt{shiny} and \texttt{golem} R packages (Chang et al., 2021; Fay et al., 2021) and is available in the \texttt{crowdforecastr} R package (Bosse, Abbott, EpiForecasts, et al., 2020b). To make a forecast in the application participants could select a predictive distribution (with the default being log-normal) to represent the probability that the forecasted quantity took certain values. Median and width of the uncertainty could be adjusted by either interacting with a figure showing their forecast or providing numerical values (see screenshot in Figure \ref{fig:screenshot} in the SI). The default shown was a repetition of the last known observation with constant uncertainty around it computed as the standard deviation of the last four changes in weekly log observed forecasts (i.e.~as \(\sigma(log(value4) - log(value3), log(value3) - log(value2), \ldots )\)).
Our interface also allowed participants to view past observations based on the hub data, as well as their forecasts, on a logarithmic scale and presented additional contextual COVID-19 data sourced from Our World in Data (2020). These data included, for example, notifications of both test positive COVID-19 cases and COVID-19 linked deaths and the number of COVID-19 tests conducted over time.

@@

We used two Bayesian semi-mechanistic models from the \texttt{EpiNow2} R package (version 1.3.3) as our model-based forecasts (Abbott, Hellewell, Hickson, et al., 2020). The first of these models, here called ``renewal model'', used the renewal equation (Fraser, 2007) to predict reported cases and deaths (see details in the SI). It estimated the effective reproduction number \(R_t\) (the average number of people each person infected at time t is expected to infect in turn) and modelled future infections as a weighted sum of past infection multiplied by \(R_t\). \(R_t\) was assumed to stay constant beyond the forecast date, roughly corresponding to continuing the latest exponential trend in infections. On the 9th of November we altered the date when \(R_t\) was assumed to be constant from two weeks prior to the date of the forecast to the forecast date, which we found to yield a more stable \(R_t\) estimate. Reported case and death notifications were obtained by convolving predicted infections over data-based delay distributions (Abbott, Hellewell, Thompson, et al., 2020; Abbott, Hellewell, Hickson, et al., 2020; epiforecasts.io/covid, 2020; Sherratt et al., 2021) to model the time between infection and report date. The renewal model was used to predict cases as well as deaths with forecasts being generated for each target separately. Death forecasts from the renewal model were therefore not informed by past cases. One submission of the renewal model on December 28th 2020 was delayed and therefore not included in the official Forecast hub ensemble.

The second model (``convolution model'', see details in SI) was only used to forecast deaths and was added later, starting December 7th 2020 (with the first forecast from December 7th suffering from a software bug and therefore disregarded in all further analyses). The convolution model was submitted, but never included in the official Forecast hub ensemble due to concerns that it could be too similar to the renewal model. The convolution model predicted deaths as a fraction of infected people who would die with some delay, by using a convolution of reported cases with a distribution that described the delay from case report to death and a scaling factor (the case-fatality ratio). Both the renewal and the convolution model used daily observations and assumed a negative binomial observation model with a multiplicative day-of-the-week effect (Abbott, Hellewell, Hickson, et al., 2020).

Line list data used to inform the prior for the delay from symptom onset to test positive case report or death in the model-based forecasts was sourced from (Xu et al., 2020) with data available up to the 1st of August. All model fitting was done using Markov-chain Monte Carlo (MCMC) in stan (Stan Development Team, 2020) with each location and forecast target being fitted separately.

@@

For the main analysis we focused mostly on two week ahead forecasts, as COVID-19 forecasts, especially for cases, were in the past found to have poor predictive performance beyond this horizon (Bracher, Wolffram, Deuschel, Görgen, Ketterer, Ullrich, Abbott, Barbarossa, Bertsimas, Bhatia, Bodych, Bosse, Burgard, Castro, et al., 2021). Forecasts for cases were scored using the full period from October 2020 until March 2021. To ensure comparability between models, all death forecasts were scored using only the period from December 14th on, where all models including the convolution model were available. To ensure robustness of our results we conducted a sensitivity analysis where all forecasts (including cases) were scored only over the later period for which all forecasts were available (see Section \ref{sensitivity-analysis} in the SI). Results remained broadly unchanged.

Forecasts were analysed using the following scoring metrics: The weighted interval score (WIS) (Bracher, Ray, et al., 2021), the absolute error, relative bias, and empirical coverage of the 50\% and 90\% prediction intervals. The WIS is a proper scoring rule (Gneiting \& Raftery, 2007), meaning that in expectation the score is optimised by reporting a predictive distribution that is identical to the true data-generating distribution. Forecasters are therefore incentivised to report their true belief about the future. The WIS can be understood as a generalisation of the absolute error to quantile-based forecasts (also meaning that smaller values are better) and can be decomposed into three separate penalties: forecast spread (i.e.~uncertainty of forecasts), over-prediction and under-prediction. While the over- and under-prediction components of the WIS capture the amount of over-prediction and under-prediction in absolute terms, we also look at a relative tendency to make biased forecasts. The bias metric (Funk et al., 2019) we use captures how much probability mass of the forecast was above or below the true value (mapped to values between -1 and 1) and therefore represents a general tendency to over- or under-predict in relative terms. A value of -1 implies that all quantiles of the predictive distribution are below the observed value and a value of 1 that all quantiles are above the observed value. Empirical coverage is the percentage of observed values that fall inside a given prediction interval (e.g.~how many observed values fall inside all 50\% prediction intervals). Scoring metrics are explained in more detail in Table \ref{tab:scoring-metrics} in the SI. All scores were calculated using the \texttt{scoringutils} R package (Bosse, Abbott, EpiForecasts, et al., 2020c).

@@

where \(I^c_t = \sum_{s< t} I_s\) are cumulative infections by \(t-1\) and \(I'_t\) are the unadjusted infections defined above. This adjustment is based on that implemented in the \texttt{epidemia} R package (Bhatt et al., n.d.; Scott et al., 2020).

>>>>

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{setspace}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Comparing human and model-based forecasts of COVID-19 in Germany and Poland},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin=1in]{geometry}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\usepackage{lineno}
\linenumbers
\usepackage{float}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi

@@

\setstretch{2}
\emph{Nikos I. Bosse, Sam Abbott, Johannes Bracher, Habakuk Hain, Billy J. Quilty, Mark Jit, Centre for the Mathematical Modelling of Infectious Diseases COVID-19 Working Group, Edwin van Leeuwen, Anne Cori, Sebastian Funk}

@@

\hypertarget{author-summary}{%
\section{Author summary}\label{author-summary}}

Mathematical models of COVID-19 have played a key role in informing governments across the world. While mathematical models are informed by our knowledge of infectious disease dynamics, they are ultimately developed and iteratively adjusted by the researchers and shaped by their subjective opinions. To investigate what modelling is able to add beyond the subjective opinion of the researcher alone, we compared human forecasts with model-based predictions of COVID-19 cases and deaths submitted to the so-called German/Polish Forecast Hub (which collates a variety of models from a range of teams).\\
We found that our human forecasts consistently outperformed an aggregate of all available model-based forecasts when predicting cases, but not when predicting deaths. Our findings suggest that human insight may be most valuable when forecasting highly uncertain quantities, which depend on many factors that are hard to model using equations, while mathematical models may be most useful in settings like predicting deaths, where leading indicators with a clear connection to the target variable are available. This potentially has very relevant policy implications, as agencies informing policy-makers could benefit from routinely eliciting human forecasts in addition to model-based predictions to inform policies.

@@

Infectious disease modelling has a long tradition and has helped inform public health decisions both through scenario modelling, as well as actual forecasts of (among others) influenza (e.g. 1,2--4), dengue fever (e.g. 5,6,7), ebola (e.g. 8,9), chikungunya (e.g. 10,11) and now COVID-19 (e.g. 12,13--17). Applications of epidemiological models differ in the way they make statements about the future. Forecasts aim to predict the future as it will occur, while scenario modelling and projections aim to represent what the future could look like under certain scenario assumptions or if conditions stayed the same as they were in the past. Forecasts can be judged by comparing them against observed data. Since it is much harder to fairly assess the accuracy and usefulness of projections and scenario modelling in the same way, this work focuses on forecasts, which represent only a subset of all epidemiological modelling.

Since March 2020, forecasts of COVID-19 from multiple teams have been collected, aggregated and compared by Forecast Hubs such as the US Forecast Hub (13,14), the German and Polish Forecast Hub (15,16) and the European Forecast Hub (17). Often, different individual forecasts are combined into a single forecast, e.g.~by taking the mean or median of all forecasts. These ensemble forecasts usually tend to perform better and more consistently than individual forecasts (see e.g. (6); (18)).

@@

Previous work has examined such direct human forecasts in various contexts, such as geopolitics (19,20), meta-science (21,22), sports (23) and epidemiology (11,24,25). Several prediction platforms (26--28) and prediction markets (29) have been created to collate expert and non-expert predictions. However, with the notable exception of (11), these forecasts were not designed to be evaluated alongside model-based forecasts and usually follow their own (often binary) prediction formats. Direct human forecasts may be able to take into account insights and relationships between variables which are hard to specify using epidemiological models. However, it is not entirely clear in which situations human forecasts perform well or badly. For example, (11) found that humans could outperform computer models at predicting the 2014/15 and 2015/16 flu season in the US, a setting where the disease was well known and information about previous seasons was available. However, humans tended to do slightly worse at predicting the 2014/15 outbreak of chikungunya in the Americas, a disease previously largely unobserved and unknown in these regions at the time.

In this study, we analyse the performance of direct human forecasts relative to model-based forecasts and discuss the added benefit of epidemiological modelling over human judgement alone. As a case study, we use different forecasts, involving varying degrees of human intervention, which we submitted in real time to the German and Polish Forecast Hub. In contrast to (11) we elicited not only point predictions, but full predictive distributions (``probabilistic forecasts'', see e.g. (30)) from participants. This allows us to compare not only predictive accuracy, but also how well human forecasters and model-based forecasts were able to quantify forecast uncertainty.

@@

We created and submitted the following forecasts to the German and Polish Forecast Hub: 1) a direct human forecast (henceforth called ``crowd forecast''), elicited from participants through a web application (31) and 2) two semi-mechanistic model-based forecasts (``renewal model'' and ``convolution model'') informed by basic assumptions about COVID-19 epidemiology. While the two semi-mechanistic forecasts were necessarily shaped by our implicit assumptions and decisions, they were designed such as to minimise the amount of human intervention involved. For example, we refrained from adjusting model outputs or refining the models based on past performance. Forecasts were created in real time over a period of 21 weeks from October 12th 2020 until March 1st 2021 and submitted to the German and Polish Forecast hub (15,16). All code and tools necessary to generate the forecasts and make a forecast submission are available in the \texttt{covid.german.forecasts} R package (32). This repository also contains a record of all forecasts submitted to the German and Polish Forecast Hub. Forecasts were evaluated using a variety of scoring metrics and compared among each other and against an ensemble of all other models submitted to the German and Polish Forecast Hub.

@@

The German and Polish Forecast Hub (now mostly merged into the (17)) elicits predictions for various COVID-19 related forecast targets from different research groups every week. Forecasts had to be made every Monday (with submissions allowed until Tuesday 3pm) and were permitted to use any data that was available by Monday 11.59pm. We submitted forecasts for incident and cumulative weekly reported numbers of cases of and deaths from COVID-19 on a national level in Germany and Poland over a one to four week forecast horizon. Forecasts were submitted on Mondays, but weeks were defined as ending on a Saturday (and starting on Sunday), meaning that forecast horizons were in fact 5, 12, 19 and 26 days. Submissions were required in a quantile-based format with 23 quantiles of each output measure at levels 0.01, 0.025, 0.05, 0.10, 0.15, \ldots, 0.95, 0.975, 0.99. Forecasts submitted to the Forecast Hub were combined into different ensembles every week, with the median ensemble (i.e., the \(\alpha\)-quantile of the ensemble is given by the median of all submitted \(\alpha\)-quantiles) being the default ensemble shown on all official Forecast hub visualisations (\url{https://kitmetricslab.github.io/forecasthub/forecast}).

Data on daily reported test positive cases and deaths linked to COVID-19 were provided by the organisers of the German and Polish Forecast hub. Until December 14th, 2020, these data were sourced from the European Centre for Disease Control (33). After ECDC stopped publishing daily data, observations were sourced from the Robert Koch Institute (RKI) and the Polish Ministry of Health for the remainder of the submission period (34). These data are subject to reporting artefacts, (such as for example delayed case reporting in Poland on the 24th November, (35)), changes in reporting over time, and variation in testing regimes (for example in Germany from the 11th of November on, (36)).

@@

Our crowd forecasts were created as an ensemble of forecasts made by individual participants every week through a web application (\url{https://cmmid-lshtm.shinyapps.io/crowd-forecast/}). Weekly forecasts had to be submitted before Tuesday 12pm every week, but participants were asked to only use any information or data that was already available by Monday night. The application was built using the \texttt{shiny} and \texttt{golem} R packages (37,38) and is available in the \texttt{crowdforecastr} R package (31). To make a forecast in the application participants could select a predictive distribution (with the default being log-normal) to represent the probability that the forecasted quantity took certain values. Median and width of the uncertainty could be adjusted by either interacting with a figure showing their forecast or providing numerical values (see screenshot in Figure \ref{fig:screenshot} in the SI). The default shown was a repetition of the last known observation with constant uncertainty around it computed as the standard deviation of the last four changes in weekly log observed forecasts (i.e.~as \(\sigma(log(value4) - log(value3), log(value3) - log(value2), \ldots )\)).
Our interface also allowed participants to view past observations based on the hub data, as well as their forecasts, on a logarithmic scale and presented additional contextual COVID-19 data sourced from (39). These data included, for example, notifications of both test positive COVID-19 cases and COVID-19 linked deaths and the number of COVID-19 tests conducted over time.

@@

We used two Bayesian semi-mechanistic models from the \texttt{EpiNow2} R package (version 1.3.3) as our model-based forecasts (40). The first of these models, here called ``renewal model'', used the renewal equation (41) to predict reported cases and deaths (see details in the SI). It estimated the effective reproduction number \(R_t\) (the average number of people each person infected at time t is expected to infect in turn) and modelled future infections as a weighted sum of past infection multiplied by \(R_t\). \(R_t\) was assumed to stay constant beyond the forecast date, roughly corresponding to continuing the latest exponential trend in infections. On the 9th of November we altered the date when \(R_t\) was assumed to be constant from two weeks prior to the date of the forecast to the forecast date, which we found to yield a more stable \(R_t\) estimate. Reported case and death notifications were obtained by convolving predicted infections over data-based delay distributions (40,42--44) to model the time between infection and report date. The renewal model was used to predict cases as well as deaths with forecasts being generated for each target separately. Death forecasts from the renewal model were therefore not informed by past cases. One submission of the renewal model on December 28th 2020 was delayed and therefore not included in the official Forecast hub ensemble.

The second model (``convolution model'', see details in SI) was only used to forecast deaths and was added later, starting December 7th 2020 (with the first forecast from December 7th suffering from a software bug and therefore disregarded in all further analyses). The convolution model was submitted, but never included in the official Forecast hub ensemble due to concerns that it could be too similar to the renewal model. The convolution model predicted deaths as a fraction of infected people who would die with some delay, by using a convolution of reported cases with a distribution that described the delay from case report to death and a scaling factor (the case-fatality ratio). Both the renewal and the convolution model used daily observations and assumed a negative binomial observation model with a multiplicative day-of-the-week effect (40).

Line list data used to inform the prior for the delay from symptom onset to test positive case report or death in the model-based forecasts was sourced from (45) with data available up to the 1st of August. All model fitting was done using Markov-chain Monte Carlo (MCMC) in stan (46) with each location and forecast target being fitted separately.

@@

For the main analysis we focused mostly on two week ahead forecasts, as COVID-19 forecasts, especially for cases, were in the past found to have poor predictive performance beyond this horizon (15). Forecasts for cases were scored using the full period from October 2020 until March 2021. To ensure comparability between models, all death forecasts were scored using only the period from December 14th on, where all models including the convolution model were available. To ensure robustness of our results we conducted a sensitivity analysis where all forecasts (including cases) were scored only over the later period for which all forecasts were available (see Section \ref{sensitivity-analysis} in the SI). Results remained broadly unchanged.

Forecasts were analysed using the following scoring metrics: The weighted interval score (WIS) (47), the absolute error, relative bias, and empirical coverage of the 50\% and 90\% prediction intervals. The WIS is a proper scoring rule (48), meaning that in expectation the score is optimised by reporting a predictive distribution that is identical to the true data-generating distribution. Forecasters are therefore incentivised to report their true belief about the future. The WIS can be understood as a generalisation of the absolute error to quantile-based forecasts (also meaning that smaller values are better) and can be decomposed into three separate penalties: forecast spread (i.e.~uncertainty of forecasts), over-prediction and under-prediction. While the over- and under-prediction components of the WIS capture the amount of over-prediction and under-prediction in absolute terms, we also look at a relative tendency to make biased forecasts. The bias metric (9) we use captures how much probability mass of the forecast was above or below the true value (mapped to values between -1 and 1) and therefore represents a general tendency to over- or under-predict in relative terms. A value of -1 implies that all quantiles of the predictive distribution are below the observed value and a value of 1 that all quantiles are above the observed value. Empirical coverage is the percentage of observed values that fall inside a given prediction interval (e.g.~how many observed values fall inside all 50\% prediction intervals). Scoring metrics are explained in more detail in Table \ref{tab:scoring-metrics} in the SI. All scores were calculated using the \texttt{scoringutils} R package (49).

@@

where \(I^c_t = \sum_{s< t} I_s\) are cumulative infections by \(t-1\) and \(I'_t\) are the unadjusted infections defined above. This adjustment is based on that implemented in the \texttt{epidemia} R package (56,57).
